Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)

### 2017-02-15 csf-2017/paper.tex
Change some \cite's to \citep's.

Compromising software security is often based on attacks that break programming
language properties relied upon by software authors, such as control flow
integrity, local state encapsulation, etc. Commodity processors offer little
support for defending against such attacks: they offer security primitives with
only coarse-grained memory protection and limited compartmentalization
scalability. As a result, defenses against attacks on control flow integrity and
local state encapsulation are either limited to only certain common forms of
attacks (leading to an attack-defense arms race) or rely on techniques like
machine code rewriting
(e.g.,~\cite{wahbe_efficient_1993,abadi_control-flow_2005}), machine code
verification (e.g.,~\cite{morrisett_system_1999}), virtual machines with a native
stack (e.g.,~the JVM~\cite{lindholm_java_2014}) or
randomization~\cite{forrest_building_1997}. The latter techniques essentially
emulate protection techniques on existing hardware, at the cost of performance,
system complexity and/or security.

\emph{Capability machines} are a type of processors, intended to
remediate these limitations by presenting a better security model at
the hardware level. They are based on old ideas
(e.g.~\cite{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999}),
but have recently received renewed interest; in particular, the CHERI
project has proposed new ideas and ways of tackling practical
challenges like backwards compatibility and realistic OS
support~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}. Capability
machines tag every word (in the register file and in memory) to
enforce a strict separation between numbers and capabilities, a kind
of pointers carrying some kind of authority. Memory capabilities carry
the authority to read and/or write to a range of memory
locations. Moreover, all capability machines
offer some form of \emph{object capabilities}, which represent the
authority to invoke a piece of code without exposing the code's
encapsulated private state (e.g., the M-Machine's enter capabilities or
CHERI's sealed code/data pairs).

>>>>

Compromising software security is often based on attacks that break programming
language properties relied upon by software authors, such as control flow
integrity, local state encapsulation, etc. Commodity processors offer little
support for defending against such attacks: they offer security primitives with
only coarse-grained memory protection and limited compartmentalization
scalability. As a result, defenses against attacks on control flow integrity and
local state encapsulation are either limited to only certain common forms of
attacks (leading to an attack-defense arms race) or rely on techniques like
machine code rewriting
(e.g.,~\cite{wahbe_efficient_1993,abadi_control-flow_2005}), machine code
verification (e.g.,~\cite{morrisett_system_1999}), virtual machines with a native
stack (e.g.,~the JVM~\citep{lindholm_java_2014}) or
randomization~\citep{forrest_building_1997}. The latter techniques essentially
emulate protection techniques on existing hardware, at the cost of performance,
system complexity and/or security.

\emph{Capability machines} are a type of processors, intended to
remediate these limitations by presenting a better security model at
the hardware level. They are based on old ideas
(e.g.~\cite{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999}),
but have recently received renewed interest; in particular, the CHERI
project has proposed new ideas and ways of tackling practical
challenges like backwards compatibility and realistic OS
support~\citep{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}. Capability
machines tag every word (in the register file and in memory) to
enforce a strict separation between numbers and capabilities, a kind
of pointers carrying some kind of authority. Memory capabilities carry
the authority to read and/or write to a range of memory
locations. Moreover, all capability machines
offer some form of \emph{object capabilities}, which represent the
authority to invoke a piece of code without exposing the code's
encapsulated private state (e.g., the M-Machine's enter capabilities or
CHERI's sealed code/data pairs).
